### 2018-05-27 paper/submit-aas/referee-response.txt
response to ref

We thank the referee for a constructive report that has improved the quality of
this work. Below, we have copied the remarks from the referee (indented by >)
and have responded to each point in text that immediately follows.

> The analysis of the Gaia DR2 data requires some care, especially at the faint
> end. For example, the authors do not discuss the amplitude of the correlations
> between the proper motion components and how these may affect their selection
> of tentative members. (By the way, the plots show relatively large ranges,
> particularly those of the proper motions and it would be useful if a few
> zoom-ins are included, so that the reader can assess better the data.).
>
> It would also be useful if the authors would explore how many stars in the
> regions around the stream have 2 parameter instead of 5 parameter solutions
> (i.e. they have position measured on the sky but no proper motions).
>
> Although the authors comment that all the stars have 8 or more visibility
> periods, this does not reflect completeness of the dataset but merely the
> reliability of the astrometric solution. Therefore maps showing the quality of
> the astrometric data in this region of the sky should be included as they
> enhance the credibility of the various features the authors claim to find. As
> a guide to what are the relevant diagnostics, the authors may wish to use
> Arenou et a. 2018 or Lindegren et al 2018.
>
> The presence of an overdensity that could correspond to the location of the
> progenitor is not so clear from the plots shown in the paper. Again a zoom-in
> might help here.

* APW:
* TODO: final figure summarize data quality and dust
* Push back on Zoom-ins? -> plot them for the referee, don't add much
* TODO: send Ana file with extinction estimates

> The authors need to include a measurement of the proper motion in the
> equatorial reference frame, for example, at the location of the highest
> density of the stream. This is clearly missing from the paper, and would make
> the paper useful to other readers.

* APW: easy, but stupid

> The authors do not use the Gaia photometry, which although shallower is likely
> to be significantly more accurate. How do the CMDs look like in the Gaia
> bands?

* AB: will check CMD
* Respond with plot and comment
* Main sequence becomes flatter

> The authors do not re-derive a distance to the stream, is there a reason for
> this? It would perhaps be useful to overlay the isochrone on the Gaia data to
> try to determine the distance to GD-1 more accurately, and hence also the
> tangential velocities of the stars in the stream.

* AB: Punt to next paper: respond that distance not the focus

> It was not clear to me how the solar reflex motion was corrected via the
> equation given, where does this stem from?

We have added a sentence to clarify how this was done.

> The significance of the various features (the blob, the spur) needs to be
> quantified. For example, Fig.4 seems to show a small overdensity in the "off
> blob" region on the proper motion diagram but there is no mention of this.

* APW/AB: Define significance

> The authors state that the proximity of the features to the location of the
> large gaps suggests that they might be related. Please elaborate on how.

* AB: explain the reference to Yoon et al., be explicit about the possible
  connection to the gap

> The orbit shown in Fig.3 is not integrated sufficiently long to be able to
> establish whether the stream has crossed the disc. Please include a comparison
> to an orbit integrated for 6-8 Gyr.

The orbit shown in Figure 3 crosses the disk midplane (z=0) twice in the segment of the orbit shown. Given uncertainties in the Milky Way's gravitational potential, and uncertainties in the full-space motion of the stream progenitor, we don't think that including a 6-8 Gyr integration in the paper will be too meaningful. However, we have extended the integration time for the paper figure to 1 Gyr, and attach a figure with a 6 Gyr integration with this response for the referee to see.

* Comments from the AAS statistics editor:

> In Fig 2, the authors choose a parabolic model for the GD-1 stream with
> (arbitrary?) lower/upper boundaries. This band clearly includes some
> background stars, and may miss some stream stars around \phi_1=-30d. It would
> be nice to identify stream members by some statistical technique, either in
> the 2D sky or 4D sky+kinematics space.

> I suggest the authors consult the excellent summary of modern `spatial point
> processes' methodology with cookbook to the comprehensive 'spatstat' R/CRAN
> package, by A.Baddeley et al. 'Spatial Point Patterns: Methodology and
> Applications with R' (2015). For example, I suspect a good list of stream
> members could emerge from a Voronoi tessellation with truncation of tile
> areas. Other possibilities include: nearest neighbors within a distance
> threshold, nonparametric density estimation, Allard-Fraley feature detection
> (MLE on Voronoi tessellation), and Byers-Raftery feature detection (MLE
> mixture model on k-NN distribution).

* APW: simple statistical technique in 4D, punt