### 2018-05-31 paper/submit-aas/referee-response.txt

We thank the referee for a constructive report that has improved the quality of
this work. Below, we have copied the remarks from the referee (indented by >)
and have responded to each point in text that immediately follows.

> The analysis of the Gaia DR2 data requires some care, especially at the faint
> end. For example, the authors do not discuss the amplitude of the correlations
> between the proper motion components and how these may affect their selection
> of tentative members. (By the way, the plots show relatively large ranges,
> particularly those of the proper motions and it would be useful if a few
> zoom-ins are included, so that the reader can assess better the data.).
>
> It would also be useful if the authors would explore how many stars in the
> regions around the stream have 2 parameter instead of 5 parameter solutions
> (i.e. they have position measured on the sky but no proper motions).
>
> Although the authors comment that all the stars have 8 or more visibility
> periods, this does not reflect completeness of the dataset but merely the
> reliability of the astrometric solution. Therefore maps showing the quality of
> the astrometric data in this region of the sky should be included as they
> enhance the credibility of the various features the authors claim to find. As
> a guide to what are the relevant diagnostics, the authors may wish to use
> Arenou et a. 2018 or Lindegren et al 2018.
>
> The presence of an overdensity that could correspond to the location of the
> progenitor is not so clear from the plots shown in the paper. Again a zoom-in
> might help here.

* APW:
* TODO: final figure summarize data quality and dust
* Push back on Zoom-ins? -> plot them for the referee, don't add much

- Dust map done
- 2-parameter solution done
- Source density at phi1 = [-40, -20, 0], phi2 = 0 is ~8000 sources/deg^2 -- should be close to 100% complete according to Arenou.
https://arxiv.org/pdf/1804.09375.pdf
https://arxiv.org/pdf/1804.09366.pdf

> The authors need to include a measurement of the proper motion in the
> equatorial reference frame, for example, at the location of the highest
> density of the stream. This is clearly missing from the paper, and would make
> the paper useful to other readers.

* APW: easy, but stupid

> The authors do not use the Gaia photometry, which although shallower is likely
> to be significantly more accurate. How do the CMDs look like in the Gaia
> bands?

The referee is correct that the Gaia photometry is significantly more accurate than PanSTARRS, however, at faint magnitudes, the regime where GD-1 members are most easily separable from the Milky Way field stars, the Gaia BP-RP color is much more uncertain than the PanSTARRS g-i color.
As a part of the response, we attach a figure that compares Gaia (left) and PanSTARRS (right) color-magnitude diagrams of GD-1 stars, selected with -1<phi_2<1, -50<phi_1<-10, and our fiducial proper motion selection.
Typical photometric uncertainties are shown as a function of magnitude, centered on 0 color in both panels.
PanSTARRS photometry is systematics dominated in the entire range, and the uncertainty in both g and i bands is 0.01 mag (0.014 mag for g-i color). 
Uncertainties in Gaia's G band are smaller in the entire range, but BP-RP's vary from 0.001 at G~15 to 0.05 at G~20, approximately 5 times larger than PanSTARRS at comparably faint magnitudes.
This leads to GD-1's main sequence being much better defined in PanSTARRS color-magnitude diagram than in Gaia's, so we opt to retain PanSTARRS photometry for selection of likely GD-1 members.

Q: do we comment on using g vs G?

> The authors do not re-derive a distance to the stream, is there a reason for
> this? It would perhaps be useful to overlay the isochrone on the Gaia data to
> try to determine the distance to GD-1 more accurately, and hence also the
> tangential velocities of the stars in the stream.

* AB: Punt to next paper: respond that distance not the focus
- we do lay an isochrone in fig 1, consistent with a previously reported distance and stellar population
- however, we note that fig 4 indicates a distance gradient, closer at trailing than leading tail, also consistent with the literature
- while re-deriving a distance to the stream using would be useful, it is rather involved and would need to be done along the stream, so we are leaving it for our next contribution where we will derive more general properties of GD-1, while focusing on the new features in this one

> It was not clear to me how the solar reflex motion was corrected via the
> equation given, where does this stem from?

We have added a sentence to clarify how this was done.

> The significance of the various features (the blob, the spur) needs to be
> quantified. For example, Fig.4 seems to show a small overdensity in the "off
> blob" region on the proper motion diagram but there is no mention of this.

* APW/AB: Define significance

> The authors state that the proximity of the features to the location of the
> large gaps suggests that they might be related. Please elaborate on how.

* AB: explain the reference to Yoon et al., be explicit about the possible
  connection to the gap

> The orbit shown in Fig.3 is not integrated sufficiently long to be able to
> establish whether the stream has crossed the disc. Please include a comparison
> to an orbit integrated for 6-8 Gyr.

The orbit shown in Figure 3 crosses the disk midplane (z=0) twice in the segment of the orbit shown. Given uncertainties in the Milky Way's gravitational potential, and uncertainties in the full-space motion of the stream progenitor, we don't think that including a 6-8 Gyr integration in the paper will be too meaningful. However, we have extended the integration time for the paper figure to 1 Gyr, and attach a figure with a 6 Gyr integration with this response for the referee to see.

* Comments from the AAS statistics editor:

> In Fig 2, the authors choose a parabolic model for the GD-1 stream with
> (arbitrary?) lower/upper boundaries. This band clearly includes some
> background stars, and may miss some stream stars around \phi_1=-30d. It would
> be nice to identify stream members by some statistical technique, either in
> the 2D sky or 4D sky+kinematics space.

> I suggest the authors consult the excellent summary of modern `spatial point
> processes' methodology with cookbook to the comprehensive 'spatstat' R/CRAN
> package, by A.Baddeley et al. 'Spatial Point Patterns: Methodology and
> Applications with R' (2015). For example, I suspect a good list of stream
> members could emerge from a Voronoi tessellation with truncation of tile
> areas. Other possibilities include: nearest neighbors within a distance
> threshold, nonparametric density estimation, Allard-Fraley feature detection
> (MLE on Voronoi tessellation), and Byers-Raftery feature detection (MLE
> mixture model on k-NN distribution).

* APW: simple statistical technique in 4D, punt