### 2018-06-03 paper/submit-aas/referee-response.txt
distance response

We thank the referee for a constructive report that has improved the quality of
this work. Below, we have copied the remarks from the referee (indented by >)
and have responded to each point in text that immediately follows.

> The analysis of the Gaia DR2 data requires some care, especially at the faint
> end. For example, the authors do not discuss the amplitude of the correlations
> between the proper motion components and how these may affect their selection
> of tentative members. (By the way, the plots show relatively large ranges,
> particularly those of the proper motions and it would be useful if a few
> zoom-ins are included, so that the reader can assess better the data.).
>
> It would also be useful if the authors would explore how many stars in the
> regions around the stream have 2 parameter instead of 5 parameter solutions
> (i.e. they have position measured on the sky but no proper motions).
>
> Although the authors comment that all the stars have 8 or more visibility
> periods, this does not reflect completeness of the dataset but merely the
> reliability of the astrometric solution. Therefore maps showing the quality of
> the astrometric data in this region of the sky should be included as they
> enhance the credibility of the various features the authors claim to find. As
> a guide to what are the relevant diagnostics, the authors may wish to use
> Arenou et a. 2018 or Lindegren et al 2018.
>
> The presence of an overdensity that could correspond to the location of the
> progenitor is not so clear from the plots shown in the paper. Again a zoom-in
> might help here.

* APW:
* TODO: final figure summarize data quality and dust
* Push back on Zoom-ins? -> plot them for the referee, don't add much

- Dust map done
- 2-parameter solution done
- Source density at phi1 = [-40, -20, 0], phi2 = 0 is ~8000 sources/deg^2 -- should be close to 100% complete according to Arenou.
https://arxiv.org/pdf/1804.09375.pdf
https://arxiv.org/pdf/1804.09366.pdf

> The authors need to include a measurement of the proper motion in the
> equatorial reference frame, for example, at the location of the highest
> density of the stream. This is clearly missing from the paper, and would make
> the paper useful to other readers.

* APW: easy, but stupid

> The authors do not use the Gaia photometry, which although shallower is likely
> to be significantly more accurate. How do the CMDs look like in the Gaia
> bands?

The referee is correct that the Gaia photometry is more accurate than PanSTARRS, however, at faint magnitudes, the regime where GD-1 members are most easily separable from the Milky Way field stars, the Gaia BP-RP color is much more uncertain than the PanSTARRS g-i color.
As a part of the response, we attach a figure that compares Gaia (left) and PanSTARRS (right) color-magnitude diagrams of GD-1 stars, selected within -1<phi_2<1, -50<phi_1<-10, and our fiducial proper motion selection.
Typical photometric uncertainties are shown as a function of magnitude, centered on 0 color in both panels.
PanSTARRS photometry is systematics dominated in the entire range, and the uncertainty in both g and i bands is 0.01 mag (0.014 mag for g-i color). 
Uncertainties in Gaia's G band are smaller in the entire range, but BP-RP's vary from 0.001 at G~15 to 0.05 at G~20, approximately 5 times larger than PanSTARRS at comparably faint magnitudes.
This leads to GD-1's main sequence being much better defined in PanSTARRS color-magnitude diagram than in Gaia's, so we opt to retain PanSTARRS photometry for selection of likely GD-1 members.

> The authors do not re-derive a distance to the stream, is there a reason for
> this? It would perhaps be useful to overlay the isochrone on the Gaia data to
> try to determine the distance to GD-1 more accurately, and hence also the
> tangential velocities of the stars in the stream.

An isochrone at a fiducial distance to the stream is already overlaid on the GD-1 color-magnitude diagram (cmd) in Figure 1, and used to define a cmd selection box, which is very useful for improving the selection of likely GD-1 members.
Since the focus of this paper is reporting on new features of GD-1 detected in a cleaner map of the stream, using literature distance to the stream as a means to improve member selection seems adequate.
Of course, more precise distances, and hence tangential velocities, would be especially useful for dynamical modeling of GD-1, but improving on the current measurements would require a more involved calculation, so we leave it for future work where we plan to comprehensively derive stream properties in all six dimensions of the phase space.

> It was not clear to me how the solar reflex motion was corrected via the
> equation given, where does this stem from?

We have added a sentence to clarify how this was done.

> The significance of the various features (the blob, the spur) needs to be
> quantified. For example, Fig.4 seems to show a small overdensity in the "off
> blob" region on the proper motion diagram but there is no mention of this.

* APW/AB: Define significance

> The authors state that the proximity of the features to the location of the
> large gaps suggests that they might be related. Please elaborate on how.

* AB: explain the reference to Yoon et al., be explicit about the possible
  connection to the gap

> The orbit shown in Fig.3 is not integrated sufficiently long to be able to
> establish whether the stream has crossed the disc. Please include a comparison
> to an orbit integrated for 6-8 Gyr.

The orbit shown in Figure 3 crosses the disk midplane (z=0) twice in the segment of the orbit shown. Given uncertainties in the Milky Way's gravitational potential, and uncertainties in the full-space motion of the stream progenitor, we don't think that including a 6-8 Gyr integration in the paper will be too meaningful. However, we have extended the integration time for the paper figure to 1 Gyr, and attach a figure with a 6 Gyr integration with this response for the referee to see.

* Comments from the AAS statistics editor:

> In Fig 2, the authors choose a parabolic model for the GD-1 stream with
> (arbitrary?) lower/upper boundaries. This band clearly includes some
> background stars, and may miss some stream stars around \phi_1=-30d. It would
> be nice to identify stream members by some statistical technique, either in
> the 2D sky or 4D sky+kinematics space.

> I suggest the authors consult the excellent summary of modern `spatial point
> processes' methodology with cookbook to the comprehensive 'spatstat' R/CRAN
> package, by A.Baddeley et al. 'Spatial Point Patterns: Methodology and
> Applications with R' (2015). For example, I suspect a good list of stream
> members could emerge from a Voronoi tessellation with truncation of tile
> areas. Other possibilities include: nearest neighbors within a distance
> threshold, nonparametric density estimation, Allard-Fraley feature detection
> (MLE on Voronoi tessellation), and Byers-Raftery feature detection (MLE
> mixture model on k-NN distribution).

* APW: simple statistical technique in 4D, punt