### 2018-06-14 paper/submit-aas/referee-response.txt
new figure and response

We thank the referee for a constructive report that has improved the quality of
this work. Below, we have copied the remarks from the referee (indented by >)
and have responded to each point in text that immediately follows.

> The analysis of the Gaia DR2 data requires some care, especially at the faint
> end. For example, the authors do not discuss the amplitude of the correlations
> between the proper motion components and how these may affect their selection
> of tentative members. (By the way, the plots show relatively large ranges,
> particularly those of the proper motions and it would be useful if a few
> zoom-ins are included, so that the reader can assess better the data.).
>
> It would also be useful if the authors would explore how many stars in the
> regions around the stream have 2 parameter instead of 5 parameter solutions
> (i.e. they have position measured on the sky but no proper motions).
>
> Although the authors comment that all the stars have 8 or more visibility
> periods, this does not reflect completeness of the dataset but merely the
> reliability of the astrometric solution. Therefore maps showing the quality of
> the astrometric data in this region of the sky should be included as they
> enhance the credibility of the various features the authors claim to find. As
> a guide to what are the relevant diagnostics, the authors may wish to use
> Arenou et a. 2018 or Lindegren et al 2018.
>
> The presence of an overdensity that could correspond to the location of the
> progenitor is not so clear from the plots shown in the paper. Again a zoom-in
> might help here.

We thank the referee for these suggestions and agree that it is useful and necessary to show the quality of the astrometric solution in this region of the sky. We have added a new Figure to the Appendix that contains maps of (1) the density of 2-parameters solutions in this region, (2) the unit weight error (as defined in Lindegren+2018, C.1), and (3) the SFD dust extinction around GD-1.

We have also computed the source density at several locations along the stream to assess whether source crowding could impact the completeness (as in Arenou+2018). We find that, at each location along the stream between phi1 = (-50, -10) the source density is never larger than ~8000 sources/deg^2 (estimated using the 2-parameter and 5-parameter sources), and is far below the densities where crowding becomes a problem.

TODO: correlations between proper motion components?

We have decided not to include further zoom-in panels, as our figures already contain many panels.

> The authors need to include a measurement of the proper motion in the
> equatorial reference frame, for example, at the location of the highest
> density of the stream. This is clearly missing from the paper, and would make
> the paper useful to other readers.

* APW: easy, but stupid

> The authors do not use the Gaia photometry, which although shallower is likely
> to be significantly more accurate. How do the CMDs look like in the Gaia
> bands?

The referee is correct that the Gaia photometry is more accurate than PanSTARRS, however, at faint magnitudes, the regime where GD-1 members are most easily separable from the Milky Way field stars, the Gaia BP-RP color is much more uncertain than the PanSTARRS g-i color.
As a part of the response, we attach a figure that compares Gaia (left) and PanSTARRS (right) color-magnitude diagrams of GD-1 stars, selected within -1<phi_2<1, -50<phi_1<-10, and our fiducial proper motion selection.
Typical photometric uncertainties are shown as a function of magnitude, centered on 0 color in both panels.
PanSTARRS photometry is systematics dominated in the entire range, and the uncertainty in both g and i bands is 0.01 mag (0.014 mag for g-i color).
Uncertainties in Gaia's G band are smaller in the entire range, but BP-RP's vary from 0.001 at G~15 to 0.05 at G~20, approximately 5 times larger than PanSTARRS at comparably faint magnitudes.
This leads to GD-1's main sequence being much better defined in PanSTARRS color-magnitude diagram than in Gaia's, so we opt to retain PanSTARRS photometry for selection of likely GD-1 members.

> The authors do not re-derive a distance to the stream, is there a reason for
> this? It would perhaps be useful to overlay the isochrone on the Gaia data to
> try to determine the distance to GD-1 more accurately, and hence also the
> tangential velocities of the stars in the stream.

An isochrone at a fiducial distance to the stream is already overlaid on the GD-1 color-magnitude diagram (cmd) in Figure 1, and used to define a cmd selection box, which is very useful for improving the selection of likely GD-1 members.
Since the focus of this paper is reporting on new features of GD-1 detected in a cleaner map of the stream, using literature distance to the stream as a means to improve member selection seems adequate.
Of course, more precise distances, and hence tangential velocities, would be especially useful for dynamical modeling of GD-1, but improving on the current measurements would require a more involved calculation, so we leave it for future work where we plan to comprehensively derive stream properties in all six dimensions of the phase space.

> It was not clear to me how the solar reflex motion was corrected via the
> equation given, where does this stem from?

We have added a sentence to clarify how this was done:
---
In detail, we project the solar velocity vector onto the tangent space at each star's sky position, convert to angular motion using the assumed distance, and subtract the solar component from each star's proper motion (this is done using the \package{astropy} and \package{gala} packages).
---

> The significance of the various features (the blob, the spur) needs to be
> quantified. For example, Fig.4 seems to show a small overdensity in the "off
> blob" region on the proper motion diagram but there is no mention of this.

* APW/AB: Define significance

> The authors state that the proximity of the features to the location of the
> large gaps suggests that they might be related. Please elaborate on how.

To clarify the possible association, we've changed this statement to:
---
Both the spur and the blob are close to large gaps in the stream, so a mechanism that displaces stars from their original orbits along the stream to orbits beyond the stream would explain both sets of features. Future observations of the full 3-D kinematics combined with detailed dynamical modeling of the GD-1 system will be able to test this scenario.
---

> The orbit shown in Fig.3 is not integrated sufficiently long to be able to
> establish whether the stream has crossed the disc. Please include a comparison
> to an orbit integrated for 6-8 Gyr.

The orbit shown in Figure 3 crosses the disk midplane (z=0) twice in the segment of the orbit shown. Given uncertainties in the Milky Way's gravitational potential, and uncertainties in the full-space motion of the stream progenitor, we don't think that including a 6-8 Gyr integration in the paper will be too meaningful. However, we have extended the integration time for the paper figure to 1 Gyr, and attach a figure with a 6 Gyr integration with this response for the referee to see.

* Comments from the AAS statistics editor:

> In Fig 2, the authors choose a parabolic model for the GD-1 stream with
> (arbitrary?) lower/upper boundaries. This band clearly includes some
> background stars, and may miss some stream stars around \phi_1=-30d. It would
> be nice to identify stream members by some statistical technique, either in
> the 2D sky or 4D sky+kinematics space.

> I suggest the authors consult the excellent summary of modern `spatial point
> processes' methodology with cookbook to the comprehensive 'spatstat' R/CRAN
> package, by A.Baddeley et al. 'Spatial Point Patterns: Methodology and
> Applications with R' (2015). For example, I suspect a good list of stream
> members could emerge from a Voronoi tessellation with truncation of tile
> areas. Other possibilities include: nearest neighbors within a distance
> threshold, nonparametric density estimation, Allard-Fraley feature detection
> (MLE on Voronoi tessellation), and Byers-Raftery feature detection (MLE
> mixture model on k-NN distribution).

The parabolic model (in phi1, phi2) is fit to the stream ridge points, then shifted up and down to define a selection region (as described in Section 3.1).

We agree that, to characterize the stream, it would be better to build a probabilistic model of the stream in all available data dimensions.
However, to be useful for future work, this would require simultaneously modeling the phase-space density of the stream so that the variations with distance and distribution of proper motions could be taken into account in a robust way.
As the focus of this work is mainly to point out unexpected, off-track features that appear related to the stream, we defer this more sophisticated modeling to future work.