### 2018-07-09 paper/submit-aas/referee-response2.txt
stuff

> However, the maps sent to the referee but not included in the paper do
> not ease my concerns on how robust their conclusions are. For example,
> the figures show that the assumption of a uniform density may be
> flawed given the number of features in the data. Especially the region
> around the spur shows a high density of large values of the unit
> weight error.
>
> Since their procedure starts with a selection of tentative members
> based on proper motions, it would be good if the authors could make a
> figure that shows both the stars selected according to the proper
> motion (i.e. the sample they already have) and the stars with 2
> parameter solutions that fall within the CMD selection. It would also
> be useful to assess the relative number of good and bad solutions in
> the blob and the spur, as well as in the regions around the gaps.
>
> The issue is not so much the quality of the data that is kept (the 5
> parameter solutions), but what does incompleteness in this dataset
> imply? The presence of stars with only 2 parameter solutions implies
> that such stars would not be selected according to the procedure
> followed by the authors, and their exclusion could potentially lead to
> "features". A plot as I propose would help to double check that the
> exclusion of such stars (if any) does not give rise to gaps or
> spurious over-densities.

TODO: Ana needs to crossmatch the 2par solutions with pan-starrs

> I would also like to ask the authors to overplot their selected GD-1
> stream stars (i.e. those shown in the bottom panel of Fig. 1), in the
> maps of the density of 2 parameter solutions, the unit weight
> error and the extinction map.

We have attached these figures to the response.

> Finally a small point regarding the presence of a progenitor. Given
> that it is very difficult to identify the feature the authors
> associated to the progenitor clearly from the figures presented, and
> its significance is not demonstrated in the paper, it does not really
> seem to deserve a mention in the abstract.

We have removed this statement from the abstract.